---
title: "Spray data - _B. glumae_"
output:
  html_document: 
    keep_md: yes
    fig_caption: yes
---

Data of 10 cultivars of rice inoculated with _B. glumae_ or mock inoculated.  Discoloration of spikelets were recorded and presented as percentage.

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(ggplot2)
```


```{r}
rice_data <- read_excel("Spray-Data-06.18.20.xlsx", 
                        col_types = c("text", "numeric", "numeric",
                                      "numeric", "numeric","numeric"))
rice_data
```

We still have to "reshape" the table to make it in longer format coding a column for treatment (Mock vs Inoculated) and temperature profile (30-22 vs 30-28).

```{r}
(rice_data_long <- rice_data %>% 
  pivot_longer(cols = c("Mock_30C-22C", "Mock_30C-28C", 
                        "Pathogen_30C-22C", "Pathogen_30C-28C"),
               names_to = "Inoculation", 
               values_to = "DiscPerc") %>%
  separate(col = Inoculation, 
            sep = "_",
            into = c("Inoculation", "TempProfile")) %>% 
  unite("Inoc_Temp", Inoculation:TempProfile, remove = FALSE))

```

# Data Exploration
 
Separating mock from pathogen inoculated:

```{r, fig.height=7, warning=FALSE}
ggplot(data = rice_data_long, aes(x = Genotype, y = DiscPerc)) +
  geom_boxplot(aes(fill = TempProfile)) +
  facet_grid(. ~ Inoculation) +
  coord_flip()
```

Looking at genotype effect:

```{r, fig.height=7, fig.width=8, warning=FALSE}
ggplot(data = rice_data_long, aes(x = Inoc_Temp, y = DiscPerc)) +
  geom_boxplot(aes(fill = TempProfile)) +
  facet_wrap(Genotype ~ ., ncol = 5) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

# Clustering analysis

```{r}
library(FactoMineR)
library(factoextra)


```

