---
title: "Spray data - B. glumae"
output:
  html_document: 
    keep_md: yes
    fig_caption: yes
editor_options: 
  chunk_output_type: inline
---

Data of 10 cultivars of rice inoculated with _B. glumae_ or mock inoculated.  Discoloration of spikelets were recorded and presented as percentage.

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(ggplot2)
```


```{r}
rice_data <- read_excel("Spray-Data-06.18.20.xlsx", 
                        col_types = c("text", "numeric", "numeric",
                                      "numeric", "numeric","numeric"))
rice_data
```

We still have to "reshape" the table to make it in longer format coding a column for treatment (Mock vs Inoculated) and temperature profile (30-22 vs 30-28).

```{r}
rice_data_long <- rice_data %>% 
  pivot_longer(cols = c("Mock_30C-22C", "Mock_30C-28C", 
                        "Pathogen_30C-22C", "Pathogen_30C-28C"),
               names_to = "Inoculation", 
               values_to = "DiscPerc") %>%
  separate(col = Inoculation, 
            sep = "_",
            into = c("Inoculation", "TempProfile")) %>% 
  unite("Inoc_Temp", Inoculation:TempProfile, remove = FALSE)

#kableExtra::kable(rice_data_long, format = "markdown")
```

# Data Exploration
 
Separating mock from pathogen inoculated:

```{r, fig.height=7, warning=FALSE}
ggplot(data = rice_data_long, aes(x = Genotype, y = DiscPerc)) +
  geom_boxplot(aes(fill = TempProfile)) +
  facet_grid(. ~ Inoculation) +
  coord_flip()
```

Looking at genotype effect:

```{r, fig.height=7, fig.width=8, warning=FALSE}
ggplot(data = rice_data_long, aes(x = Inoc_Temp, y = DiscPerc)) +
  geom_boxplot(aes(fill = TempProfile)) +
  scale_fill_manual(values = alpha(c("#01665e","#8c510a"), .7)) +
  facet_wrap(Genotype ~ ., ncol = 5) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

# Clustering analysis

Since we are dealing with continuous data on four different conditions with need to scale them to estimate their relationships.

## Kmeans

```{r}
library(FactoMineR)
library(factoextra)
library(RColorBrewer)
```


```{r}
#Need to remove NAs and calculate means
rice_data_NoNAs <- rice_data_long %>% 
  drop_na() %>%
  group_by(Genotype, Inoc_Temp) %>%
  summarise(meanDiscPerc = mean(DiscPerc))

rice_matrix <- rice_data_NoNAs %>% 
  pivot_wider(names_from = "Inoc_Temp", 
              values_from = "meanDiscPerc") %>%
  column_to_rownames(var = "Genotype")

# rice_matrix_all <- rice_data_long %>%
#   drop_na() %>%
#   unite(gen_rep, Genotype, Rep, sep = "_", remove = FALSE) %>%
#   select(-c(Genotype, Rep, Inoculation, TempProfile)) %>%
#   pivot_wider(names_from = "Inoc_Temp",
#               values_from = "DiscPerc") %>%
#   column_to_rownames(var = "gen_rep")
# 
# rice_matrix_all <- na.omit(rice_matrix_all)
```


```{r}
set.seed(123)
(cluster_number <- NbClust::NbClust(rice_matrix, distance = "euclidean", min.nc = 2, 
                                   max.nc = 10, method = "complete", index = "all"))
```


```{r}
#K-means
set.seed(123)
km.res4 <- kmeans(scale(rice_matrix), 4, nstart = 10)

fviz_cluster(km.res4, data = rice_matrix, labelsize = 10)
```




```{r, fig.width=10}
#Data Clusters
#rice_Kcluster <- cbind(rice_data_NoNAs, km.res3$cluster)

rice_Kcluster_data <- drop_na(rice_data) %>%
  pivot_longer(cols = c("Mock_30C-22C", "Mock_30C-28C", 
                        "Pathogen_30C-22C", "Pathogen_30C-28C"),
               names_to = "Inoculation", 
               values_to = "DiscPerc") %>%
  separate(col = Inoculation, 
            sep = "_",
            into = c("Inoculation", "TempProfile")) %>% 
  unite("Inoc_Temp", Inoculation:TempProfile, remove = FALSE) %>%
  left_join(as_tibble(km.res4$cluster, rownames="Genotype"), 
            by = "Genotype") %>%
  rename(cluster = "value") %>%
  mutate(response = recode(cluster, '1'='Temperature dependent', 
                                    '2'='Temperature dependent', 
                                    '3'='Temperature independent', 
                                    '4'='Temperature independent')) %>%
  mutate(response_dis = recode(cluster, '1'='Disease severity (>50%)', 
                                    '2'='Disease severity (<50%)', 
                                    '3'='Disease severity (>50%)', 
                                    '4'='Disease severity (<50%)'))
  

ggplot(rice_Kcluster_data, aes(x = as.factor(cluster), y = DiscPerc)) + 
  geom_boxplot(aes(fill = Inoc_Temp)) +
  facet_wrap(vars(response, response_dis),nrow = 1, scales = "free_x") +
  scale_fill_manual(values = alpha(c("#80cdc1","#01665e","#dfc27d","#8c510a"), .7)) +
  labs(title = "K-means clustering", x="Cluster", 
       y="Discoloration Percent (%)", 
       fill="Inoculation & \nTemperature profile")
```

```{r}
(Kmeans_summary <- rice_Kcluster_data %>%
  group_by(Genotype, Inoc_Temp, response, response_dis, cluster) %>%
  summarise(meanDiscPerc = mean(DiscPerc)) %>%
  pivot_wider(names_from = Inoc_Temp,
              values_from = meanDiscPerc) %>%
  arrange(cluster)
)

```


## Hierarchical clustering
```{r, message=FALSE, fig.width=10, fig.height=5}
#Clustering
rice_hc <- hcut(rice_matrix, 4,stand = T, hc_method = "ward.D", hc_metric = "euclidean")

#checking clusters
fviz_cluster(rice_hc)
fviz_silhouette(rice_hc, print.summary = F)

#Graphical view
p <- fviz_dend(rice_hc, rect = T, cex=0.8, horiz = T, repel = TRUE, color_labels_by_k = T, 
                k_colors = "jco", rect_fill = T, rect_border = "jco")
(p2 <- p + annotate(x=-2, xend=-2, y=0, yend=12, colour="dark gray", lwd=0.5, geom="segment"))
    
#Data Clusters
rice_cluster <- cbind(rice_matrix, rice_hc$cluster)

rice_cluster_data <- rice_cluster %>%
  rownames_to_column(var = "Genotype") %>%
    pivot_longer(cols = c("Mock_30C-22C", "Mock_30C-28C", 
                        "Pathogen_30C-22C", "Pathogen_30C-28C"),
               names_to = "Inoculation", 
               values_to = "DiscPerc") %>%
  separate(col = Inoculation, 
            sep = "_",
            into = c("Inoculation", "TempProfile")) %>% 
  unite("Inoc_Temp", Inoculation:TempProfile, remove = FALSE) %>%
  dplyr::rename(cluster = "rice_hc$cluster") %>%
  mutate(response = recode(cluster, '1'='Temperature independent', 
                                    '2'='Temperature dependent', 
                                    '3'='Temperature independent', 
                                    '4'='Temperature dependent')) %>%
  mutate(response_dis = recode(cluster, '1'='Disease severity (<50%)', 
                                    '2'='Disease severity (>50%)', 
                                    '3'='Disease severity (>50%)', 
                                    '4'='Disease severity (<50%)'))

ggplot(rice_cluster_data, aes(x = as.factor(cluster), y = DiscPerc)) +   geom_boxplot(aes(fill = Inoc_Temp)) +
  facet_wrap(vars(response, response_dis),nrow = 1, scales = "free_x") +
  scale_fill_manual(values = alpha(c("#80cdc1","#01665e","#dfc27d","#8c510a"), .7)) +
  labs(x= "Cluster", y = "% Discolored spikelets", fill="Treatment") +
  theme(text = element_text(size=12))
```


```{r}
(hc_summary <- rice_cluster_data %>%
  group_by(Genotype, Inoc_Temp, response, response_dis, cluster) %>%
  summarise(meanDiscPerc = mean(DiscPerc)) %>%
  pivot_wider(names_from = Inoc_Temp,
              values_from = meanDiscPerc) %>%
  arrange(cluster)
)
```
### Comparing hierarchical clusters and kmean clustering

```{r}
(Comparison_genotypes <- full_join(Kmeans_summary, hc_summary, by = "Genotype") %>%
  rename_with(~gsub("\\.x", "\\.kmeans", .x)) %>%
  rename_with(~gsub("\\.y", "\\.hc", .x))
)

library(kableExtra)
kbl(Comparison_genotypes) %>%
   kable_styling(bootstrap_options = c("striped", "hover"))
write_csv(Comparison_genotypes, "Comparison_genotypes.csv")
```


## PCA

This first one is using FactoMiner, 
```{r}
#Need to remove NAs
rice_data_NoNAs <- na.omit(rice_data)

#Creating Matrix for analysis
rice_matrix <- rice_data_NoNAs[,3:6]
row.names(rice_matrix) <- paste0(rice_data_NoNAs$Genotype,"_",rice_data_NoNAs$Rep)

#PCA
rice_PCA <- prcomp(rice_matrix, center = T, scale. = T)


#Since we have 20 cultivars we need a palette with 20 colors
colors_n <- length(unique(rice_data_NoNAs$Genotype))
getPalette <- colorRampPalette(brewer.pal(9, "Dark2"))

(PCA_rice <- fviz_pca_biplot(rice_PCA, col.var = "blue",
                label = "var", repel = T,
                habillage = rice_data_NoNAs$Genotype, addEllipses = TRUE, ellipse.level = 0.95,
             ellipse.type ="confidence") +
  scale_color_manual(values = getPalette(colors_n)) +
  scale_shape_manual(values = c(rep(19, 5), rep(16,5), rep(17,5), rep(18,5))))

plotly::ggplotly(PCA_rice)
```

```{r}
library(viridis)
#PCA
rice_PCA <- prcomp(rice_matrix, center = T, scale. = T)

#PCA
biplot <- ggbiplot::ggbiplot(rice_PCA, obs.scale = 1, var.scale = 1) +
  geom_text(aes(label=rice_data_NoNAs$Genotype), size = 2, nudge_y = 0.2, alpha=0.5) +
  geom_point(aes(colour=rice_data_NoNAs$`Pathogen_30C-22C`)) +
  scale_color_viridis(name = "Pathogen_30C-22C", option = "D")

plotly::ggplotly(biplot)
```

```{r}
(final_data <- rice_Kcluster_data %>%
  filter(Genotype %in% c("311078","311677","310645","310998","311735","311383", "311795","310131","310802"))
)

rice.lm <- lm(DiscPerc ~ Inoculation*TempProfile*Genotype, data = final_data)

car::Anova(rice.lm, type ="III")

emmeans::emmip(rice.lm, Genotype ~ TempProfile | Inoculation)

final_data %>%
  group_by(Genotype) %>%
  rstatix::anova_test(DiscPerc ~ Inoculation*TempProfile, error = rice.lm)
```



```{r}
sessionInfo()
```

