---
title: "Spray data - _B. glumae_"
output:
  html_document: 
    keep_md: yes
    fig_caption: yes
editor_options: 
  chunk_output_type: inline
---

Data of 10 cultivars of rice inoculated with _B. glumae_ or mock inoculated.  Discoloration of spikelets were recorded and presented as percentage.

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(ggplot2)
```


```{r}
rice_data <- read_excel("Spray-Data-06.18.20.xlsx", 
                        col_types = c("text", "numeric", "numeric",
                                      "numeric", "numeric","numeric"))
rice_data
```

We still have to "reshape" the table to make it in longer format coding a column for treatment (Mock vs Inoculated) and temperature profile (30-22 vs 30-28).

```{r}
rice_data_long <- rice_data %>% 
  pivot_longer(cols = c("Mock_30C-22C", "Mock_30C-28C", 
                        "Pathogen_30C-22C", "Pathogen_30C-28C"),
               names_to = "Inoculation", 
               values_to = "DiscPerc") %>%
  separate(col = Inoculation, 
            sep = "_",
            into = c("Inoculation", "TempProfile")) %>% 
  unite("Inoc_Temp", Inoculation:TempProfile, remove = FALSE)

#kableExtra::kable(rice_data_long, format = "markdown")
```

# Data Exploration
 
Separating mock from pathogen inoculated:

```{r, fig.height=7, warning=FALSE}
ggplot(data = rice_data_long, aes(x = Genotype, y = DiscPerc)) +
  geom_boxplot(aes(fill = TempProfile)) +
  facet_grid(. ~ Inoculation) +
  coord_flip()
```

Looking at genotype effect:

```{r, fig.height=7, fig.width=8, warning=FALSE}
ggplot(data = rice_data_long, aes(x = Inoc_Temp, y = DiscPerc)) +
  geom_boxplot(aes(fill = TempProfile)) +
  facet_wrap(Genotype ~ ., ncol = 5) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

# Clustering analysis

Since we are dealing with continous data on four different conditions with need to scale them to estimate
their coordinates.
```{r, message=FALSE}
rice_wider <- rice_data %>% pivot_wider(names_from = Rep, names_sep = "_", 
                          values_from = c("Mock_30C-22C", "Mock_30C-28C", 
                          "Pathogen_30C-22C", "Pathogen_30C-28C")) %>%
  column_to_rownames("Genotype")

rice_wider_NoNAs <- rice_data %>% 
  pivot_wider(names_from = Rep, names_sep = "_", 
                          values_from = c("Mock_30C-22C", "Mock_30C-28C", 
                          "Pathogen_30C-22C", "Pathogen_30C-28C")) %>%
  column_to_rownames("Genotype") %>%
  filter(!is.na)
  

rice_scaled <- scale(rice_wider)


#kableExtra::kable(rice_scaled, format = "markdown")

#Need to remove NAs
rice_scaled_NoNAs <- na.omit(rice_scaled)

#Calculating euclidean distances
rice_dist <- dist(rice_scaled, method = "euclidean")

#Hierarchical clustering
rice_clust <- hclust(rice_scaled, method = "complete")

plot(rice_clust)
```

## PCA

```{r}
library(FactoMineR)
library(factoextra)

rice_PCA <- PCA(rice_wider, scale.unit = TRUE, graph = T)

rice_PCA$eig

fviz_pca_biplot(rice_PCA, label = )
```

